\large Neural and Symbolic Optimization of Cosmological Particle-Mesh Simulation \vspace{0.3em}

### 2024-03-27 docs/papers/sto/so.tex
FIXUP draft minor

@@

For both the $f$ and $g$ spatial optimization NNs (SONNs), we adopt
multilayer perceptron (MLP, i.e. fully connected) structures to learn
the nonlinear functions.
Both $f$ and $g$ networks have $XXX$ hidden layers with $XXX\times n_{\rm
input}$ nodes for each layer, where $n_{\rm input}$ is the size of the
input layer, which is [...] and [...] for $f$ and $g$ respectively.
The rectified linear unit (ReLU) activation function is used for hidden
layers.
We further apply the exponential function to the output as a regulator
for positive output.
Being embedded in the differentiable simulation library \pmwd, the NNs
can be trained using gradient-based optimization methods as in usual
deep learning tasks.
The parameters of the NNs, which enter the force evaluation in every
time integration step, can be treated just the same as other
cosmological parameters.
The objective gradients are accumulated, based on which the parameters
are then optimized.

With discrete Fourier transform (DFT) on a regular grid of length $N_c$
and cell size $l_c$, the range of sampled 1D discrete wavenumbers is
limited by the fundamental and Nyquist values, i.e. $k_{\rm fund} =
2\pi/(l_c N_c)$ and $k_{\rm Nyquist} = \pi/l_c$, where $l_c N_c$ is
simply the box size.
Then the normzalized wavenumber $k/k_{\rm Nyquist}$ gives the same range
$[2/N_c, 1]$ for different box sizes.
For the 3D wavenumber $k=\sqrt{\sum_{i=1}^{3}k_i^2}$, the maximum
normalized value is then $\sqrt{3}$.

@@

To make the SO more interpretable and applicable to other simulations,
we fit the trained NNs with symbolic regression for analytic
expressions of the $f(k_i; \vvartheta)$ and $g(k_1, k_2, k_3;
\vvartheta)$ functions.

@@

In \autoref{fig:tfcc}, we show the results of TF and CC for both
training and test data samples.
To show different box sizes, we normalize the wavenumber by the Nyquist
wavenumber $k_{\rm Nyquist} = \pi / l_p$ of the particle grid, where
$l_p$ is the particle spacing.
Compared to original PM, the TF of PM-SONN is improved to be consistent
with \GADGET\ within 10\% up to the particle grid Nyquist wavenumber.

@@

\vspace{1em}
\subsection{Input Features}
\label{sec:features}

>>>>

\subsection{Target Simulation and Data}

@@

For both the $f$ and $g$ spatial optimization NNs (SONNs), we adopt
multilayer perceptron (MLP, i.e. NN with fully connected layers)
structures to learn the nonlinear functions.
Both $f$ and $g$ networks have $XXX$ hidden layers with $XXX\times n_{\rm
input}$ nodes for each layer, where $n_{\rm input}$ is the size of the
input layer, which is [...] and [...] for $f$ and $g$ respectively.
The rectified linear unit (ReLU) activation function is used for hidden
layers.
We further apply the exponential function to the output as a regulator
for positive output.
Being embedded in the differentiable simulation library \pmwd, the NNs
can be trained using gradient-based optimization methods as in usual
deep learning tasks.
The parameters of the NNs, which enter the force evaluation in every
time integration step, can be treated just the same as other
cosmological parameters.
The objective gradients are accumulated, based on which the parameters
are then optimized.

@@

To make the SO more interpretable and applicable to other simulations,
we fit the trained NNs with SR for analytic expressions of the $f(k_i;
\vvartheta)$ and $g(k_1, k_2, k_3; \vvartheta)$ functions.

The data for SR fitting is generated from the trained NNs, while the
input features for SR could be different from those of the NNs.
Based on the NN input features, we could further design other features
for SR fitting.

@@

In \autoref{fig:tfcc}, we show the results of TF and CC for both
training and test data samples.
To show different box sizes, we normalize the wavenumber by the Nyquist
wavenumber $k_{\rm Nyquist} = \pi / l_p$ of the particle grid, where
$l_p$ is the particle spacing.
With discrete Fourier transform (DFT) on a regular grid of length $N_c$
and cell size $l_c$, the range of sampled 1D discrete wavenumbers is
limited by the fundamental and Nyquist values, i.e. $k_{\rm fund} =
2\pi/(l_c N_c)$ and $k_{\rm Nyquist} = \pi/l_c$, where $l_c N_c$ is
simply the box size.
Then the normzalized wavenumber $k/k_{\rm Nyquist}$ gives the same range
$[2/N_c, 1]$ for different box sizes.
For the 3D wavenumber $k=\sqrt{\sum_{i=1}^{3}k_i^2}$, the maximum
normalized value is then $\sqrt{3}$.
Compared to original PM, the TF of PM-SONN is improved to be consistent
with \GADGET\ within 10\% up to the particle grid Nyquist wavenumber.

@@

\vspace{1em}
\subsection{Physical Features}
\label{sec:features}
